---
title: "Premier League Transfer Data"
author: "Clare Clingain"
date: "August 10, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rvest)
library(dplyr)
library(tidyr)
library(stringr)
```


```{r}
url <- read_html('https://www.transfermarkt.co.uk/premier-league/neuimland/wettbewerb/GB1/saison_id/alle/land_id/alle/ausrichtung//spielerposition_id/alle/altersklasse//w_s//plus/1')
```

```{r}
url %>%
  html_nodes(".page") %>%
  html_text() %>%
  as.numeric()
```
There are two classes we need -- odd and even.

```{r}
oddrows <- read_html(url) %>%
  html_nodes(".odd") %>%
  html_text() 

evenrows <- read_html(url) %>%
  html_nodes(".even") %>%
  html_text()
```

We know there are 21 pages that follow this sort of url: https://www.transfermarkt.co.uk/premier-league/neuimland/wettbewerb/GB1/ajax/yw1/saison_id/alle/land_id/alle/ausrichtung//spielerposition_id/alle/altersklasse//w_s//plus/page/1//page/14

Let's loop through each page to get the data.

```{r}
# Get the URLS
urls <- NULL
for (i in 1:21){
  urls[i] <- paste0('https://www.transfermarkt.co.uk/premier-league/neuimland/wettbewerb/GB1/saison_id/alle/land_id/alle/ausrichtung//spielerposition_id/alle/altersklasse//w_s//plus/page/1//page/',i, sep = "")
  }
```

We will get the odd and even row separately since they are under different nodes.

```{r}
getodddata <- function(url) {
oddrows <- read_html(url) %>% 
  html_nodes(".odd") %>% 
  html_text() %>% 
  gsub('[\r\n\t]', '', .)
}

getevendata <- function(url) {
  evenrows <- read_html(url) %>% 
    html_nodes(".even") %>% 
    html_text() %>% 
    gsub('[\r\n\t]', '', .)
}

myodddata <- lapply(urls,getodddata)
myevendata <- lapply(urls, getevendata)
```

Next, we will combine the data so it'll be easier to parse afterwards as a whole.

```{r}
mydata1 <- do.call(c,myevendata)
mydata2 <- do.call(c,myodddata)
mydata <- as.data.frame(c(mydata1,mydata2), stringsAsFactors = F)
colnames(mydata) <- "player"
```

Now that we have a data frame set up, we can start to declutter the text into meaningful columns of data.

# Data Cleaning

## Extract Position

```{r}
positions <- c("Right Winger","Left-Back","Left Winger", "Centre-Back","Centre-Forward","Attacking Midfield","Right-Back","Defensive Midfield","Central Midfield", "Left Winger", "GoalKeeper","Second Striker","Left Midfield","Right Midfield")

mydata <- mydata %>%
            mutate(position = str_extract(mydata$player, "Right Winger|Left-Back|LeftWinger|Centre-Back|Centre-Forward|Attacking Midfield|Right-Back|Defensive Midfield|Central Midfield|Left Winger|Goalkeeper|Second Striker|Left Midfield|Right Midfield"))


#Remove the instances of these positions
mydata2 <- mydata %>% mutate(player = str_remove(mydata$player, "Right Winger|Left-Back|LeftWinger|Centre-Back|Centre-Forward|Attacking Midfield|Right-Back|Defensive Midfield|Central Midfield|Left Winger|Goalkeeper|Second Striker|Left Midfield|Right Midfield")) 

```

## Extract Season

This is also another element that we can easily get from the text (before we have to do anything a bit whacky).

```{r}
mydata2 <- mydata2 %>% mutate(season = str_extract(mydata2$player,"[0-9][0-9]/[0-9][0-9]"))
# Remove for clarity
mydata2$player <-  str_remove(mydata2$player,"[0-9][0-9]/[0-9][0-9]")
```

## Extract fee + market value

We're not really interested in using this part of the data, so this solution may seem a bit unsatisfactory to those who are interested. This doesn't capture the players who had a market value of less than £1,000,000. These can be recovered and updated. Some players ended up leaving on a free transfer, and therefore end up having both fee and market value equal to the same number. 

```{r}
 #str_extract(mydata2$player,"£[0-9]+[a-z]+") # test
 #str_extract(mydata2$player,"£[0-9]+\\.[0-9]+[a-z]+") # market value
 #str_extract_all(mydata2$player,"£[0-9]+\\.[0-9]+[a-z]+") # market value and fee

 
 
 temp2 <-  str_extract_all(mydata2$player,"£[0-9]+\\.[0-9]+[a-z]+")
 temp3 <- do.call(rbind,temp2)
 
 mydata3 <- mydata2 %>% 
              mutate(marketvalue = temp3[,1]) %>%
              mutate(fee = temp3[,2])
 # Remove the fee information from the first column
 mydata3 <- mydata3 %>% mutate(player = str_remove_all(mydata2$player,"£[0-9]+\\.[0-9]+[a-z]+"))
 mydata3 <- mydata3 %>%mutate(player = str_remove_all(mydata3$player,"£"))
```

## Remove Row Number

```{r}
mydata3 <- mydata3 %>% mutate(player = str_remove(mydata3$player, "[0-9]+"))
```


## Extract the Transfer date

```{r}
mydata3 <- mydata3 %>% 
 mutate(date = str_extract(mydata3$player,"[A-z][a-z]+[:space:][0-9]+\\,[:space:][0-9]+")) %>% mutate(player = str_remove(mydata3$player,"[A-z][a-z]+[:space:][0-9]+\\,[:space:][0-9]+"))
```

## Extract Player's Age

```{r}
mydata3 <- mydata3 %>% mutate(age = str_extract(mydata3$player,"[0-9]+")) %>%
  mutate(player = str_remove(mydata3$player, "[0-9]+"))
```

## Replace Accented Letters

```{r}
mydata3 <- mydata3 %>% mutate(player = str_replace_all(mydata3$player,"Á","A"))
mydata3 <- mydata3 %>% mutate(player = str_replace_all(mydata3$player,"á","a"))
mydata3 <- mydata3 %>% mutate(player = str_replace_all(mydata3$player,"í","i"))
mydata3 <- mydata3 %>% mutate(player = str_replace_all(mydata3$player,"é","e"))
mydata3 <- mydata3 %>% mutate(player = str_replace_all(mydata3$player,"É","E"))
mydata3 <- mydata3 %>% mutate(player = str_replace_all(mydata3$player,"ü","u"))
mydata3 <- mydata3 %>% mutate(player = str_replace_all(mydata3$player,"ã","a"))
mydata3 <- mydata3 %>% mutate(player = str_replace_all(mydata3$player,"ú","u"))
mydata3 <- mydata3 %>% mutate(player = str_replace_all(mydata3$player,"ö","o"))
mydata3 <- mydata3 %>% mutate(player = str_replace_all(mydata3$player,"Ö","O"))
mydata3 <- mydata3 %>% mutate(player = str_replace_all(mydata3$player,"ó","o"))
mydata3 <- mydata3 %>% mutate(player = str_replace_all(mydata3$player,"è","e"))
mydata3 <- mydata3 %>% mutate(player = str_replace_all(mydata3$player,"ë","e"))
mydata3 <- mydata3 %>% mutate(player = str_replace_all(mydata3$player,"ï","i"))
mydata3 <- mydata3 %>% mutate(player = str_replace_all(mydata3$player,"ñ","n"))

```

# Extract the League (to and from)

```{r}
league <- str_extract_all(mydata3$player, "Ligue 1|Premier League|Serie A|LaLiga|La Liga|Bundesliga|Liga NOS|Eredivisie|Super Lig|Premier Liga|LynTippeligaen|Vyshcha Liga|Super League|Championship|Premiership|Jupiler Pro League|PTT Telecomp|Argentina|HNL|Primera Division|SuperLiga|Scotland|Serie B|Superligaen Championship|Superligaen|Croatia|Uruguay|Qatar|Liga MX Clausura|Brazil|B2ª B - Grupo III|Liga 1|Ligue 2|Without Club|Torneo Final")
league2 <- do.call(rbind,league)
mydata4 <- mydata3 %>% mutate(LeagueFrom = league2[,1]) %>%
  mutate(LeagueTo = league2[,2]) %>% mutate(player = str_remove_all(mydata3$player,"Ligue 1|Premier League|Serie A|LaLiga|La Liga|Bundesliga|Liga NOS|Eredivisie|Super Lig|Premier Liga|LynTippeligaen|Vyshcha Liga|Super League|Championship|Premiership|Jupiler Pro League|PTT Telecomp|Argentina|HNL|Primera Division|SuperLiga|Scotland|Serie B|Superligaen Championship|Superligaen|Croatia|Uruguay|Qatar|Liga MX Clausura|Brazil|B2ª B - Grupo III|Liga 1|Ligue 2|Without Club|Torneo Final" ))
```

# Extract Name and Team

```{r}
# Clean up data a little more
mydata4$player <- str_replace_all(mydata4$player,"[:punct:]|[0-9]|[0-9]+[a-z]|[0-9]|Loan fee|fee|<U+200E>","")


temp <- str_extract_all(mydata4$player,"[A-Z]+[a-z]+")
```
