---
title: "Premier League Transfer Data"
author: "Clare Clingain"
date: "August 10, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rvest)
library(dplyr)
```


```{r}
url <- read_html('https://www.transfermarkt.co.uk/premier-league/neuimland/wettbewerb/GB1/saison_id/alle/land_id/alle/ausrichtung//spielerposition_id/alle/altersklasse//w_s//plus/1')
```

```{r}
url %>%
  html_nodes(".page") %>%
  html_text() %>%
  as.numeric()
```
There are two classes we need -- odd and even.

```{r}
oddrows <- read_html(url) %>%
  html_nodes(".odd") %>%
  html_text() 

evenrows <- read_html(url) %>%
  html_nodes(".even") %>%
  html_text()
```

We know there are 21 pages that follow this sort of url: https://www.transfermarkt.co.uk/premier-league/neuimland/wettbewerb/GB1/ajax/yw1/saison_id/alle/land_id/alle/ausrichtung//spielerposition_id/alle/altersklasse//w_s//plus/page/1//page/14

Let's loop through each page to get the data.

```{r}
# Get the URLS
urls <- NULL
for (i in 1:21){
  urls[i] <- paste0('https://www.transfermarkt.co.uk/premier-league/neuimland/wettbewerb/GB1/saison_id/alle/land_id/alle/ausrichtung//spielerposition_id/alle/altersklasse//w_s//plus/page/1//page/',i, sep = "")
  }
```

We will get the odd and even row separately since they are under different nodes.

```{r}
getodddata <- function(url) {
oddrows <- read_html(url) %>% 
  html_nodes(".odd") %>% 
  html_text() %>% 
  gsub('[\r\n\t]', ' ', .)
}

getevendata <- function(url) {
  evenrows <- read_html(url) %>% 
    html_nodes(".even") %>% 
    html_text() %>% 
    gsub('[\r\n\t]', ' ', .)
}

myodddata <- lapply(urls,getodddata)
myevendata <- lapply(urls, getevendata)
```

Next, we will combine the data so it'll be easier to parse afterwards as a whole.

```{r}
mydata1 <- do.call(c,myevendata)
mydata2 <- do.call(c,myodddata)
mydata <- as.data.frame(c(mydata1,mydata2), stringsAsFactors = F)
```

Now that we have a data frame set up, we can start to declutter the text into meaningful columns of data.

```{r}

```

