---
title: "Evolution of the Top 4/6"
author: "Clare Clingain"
date: "May 15, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(foreign)
library(dplyr)
require(mclust)
library(NbClust)
require(phyclust)
library(cluster)
library(rgl)
library(e1071)
EPL_aggregate <- read.csv("C:/Users/Clare/Documents/EPL Data/Clean Data/EPL_aggregate.csv")
```

We can only use teams that have been consistently in the Premier League. The teams who have accomplished this feat are Arsenal, Chelsea, Everton, Liverpool, Manchester United, and Tottenham.

```{r}
AllLeague <- EPL_aggregate %>% filter(Team == "Arsenal" | Team == "Chelsea" | Team == "Everton" |
                                        Team == "Liverpool" | Team == "Man United" |
                                        Team == "Tottenham")
#AllLeague <- AllLeague[,-c(1:2,5:6)]
AllLeague  <- AllLeague %>% filter(Week==38) %>% group_by(Team) %>%
  mutate(YellowCardSum = sum(Yellow_Cards)) %>%
  mutate(RedCardSum = sum(Red_Cards)) %>% 
  mutate(GoalsSum = sum(Goals_Scored)) %>%
  mutate(GoalsAgainstSum = sum(Goals_Conceded)) %>% 
  mutate(FoulsSum = sum(Fouls_Committed)) %>% 
  mutate(FoulsAgainstSum = sum(Fouls_Against)) %>%
  mutate(CornersSum = sum(Corners)) %>% 
  mutate(CornersConcededSum = sum(Corners_Conceded)) %>% 
  mutate(ShotsSum = sum(Shots_Taken)) %>%
  mutate(ShotsTargetSum = sum(Shots_on_Target)) %>%
  mutate(ShotsConcededSum = sum(Shots_Conceded)) %>% 
  mutate(ShotsTargetConSum = sum(Shots_on_Target_Conceded))

AllLeague <- AllLeague[c(1:4, 27:38)]
```

##Standardize the data

```{r}
AllLeague_std<- AllLeague
AllLeague_std[,-(1:4)] <- scale(AllLeague_std[,-(1:4)])
pca<- princomp(AllLeague_std[,-(1:4)])
summary(pca)
```

##Screeplot

```{r}
screeplot(pca,type="l", main = "Screeplot of PCA")
```

##Eigenvalues

```{r}
eigen(cor(AllLeague_std[,-(1:4)]))$values 

```
5 of them are greater than 1. Seems safer to go with 4 components. 

##Loadings

```{r}
pca$loadings
```

##Initial Visualization

```{r}
plot(pca$scores[,1:2],col=1)
text(pca$scores[,1:2],col=1, labels=AllLeague$Team,cex=.5)
#3D plot
plot3d(pca$scores[,1:3], cex=1.5)
text3d(pca$scores[,1:3], text=AllLeague$Team, cex=.5)
```

##Clustering

###2-cluster solution

```{r}
set.seed(2011)
km.2pca <- kmeans(pca$scores, 2, nstart= 100)
c.crit <- function(km.obj) {
  #based on k-means, for convenience due to amt of addl info in the km result object.
  #cd be generalized.
  sizes <- km.obj$size
  n <- sum(sizes)
  g <- length(sizes)
  msW<-sum(km.obj$withinss)/(n-g)
  overall.mean <- apply(km.obj$centers*km.obj$size,2,sum)/sum(km.obj$size)
  msB<-sum(km.obj$size*(t(t(km.obj$centers)-overall.mean))^2)/(g-1)
  list(msB=msB,msW=msW,C.g=msB/msW)
}
c.crit(km.2pca)$C.g
```
###3-cluster solution

```{r}
set.seed(2011)
km.3pca <- kmeans(pca$scores, 3, nstart= 100)
c.crit(km.3pca)$C.g
```
###4-cluster solution

```{r}
set.seed(2011)
km.4pca <- kmeans(pca$scores, 4, nstart= 100)
c.crit(km.4pca)$C.g
```

###5-cluster solution

```{r}
set.seed(2011)
km.5pca <- kmeans(pca$scores, 5, nstart= 100)
c.crit(km.5pca)$C.g
```
###6-cluster solution

```{r}
set.seed(2011)
km.6pca <- kmeans(pca$scores, 6, nstart= 100)
c.crit(km.6pca)$C.g
```
As such, each team belongs in its own cluster. 

##Updated Visualization

```{r}
plot3d(pca$scores)
text3d(pca$scores, text =AllLeague$Team, col = km.6pca$clust, cex=.7,pos=3)
```

##Mclust

```{r}
mcl <- Mclust(AllLeague_std[c(5:16)])
summary(mcl)
```
##NbClust

Does not run because of small sample. 

```{r}
#k.means <- NbClust(AllLeague_std[c(5:16)],method='kmeans',index='ch')
#k.means$Best.nc
```

##Apply to all teams

