---
title: "Aggregating Data"
output: github_document
---


### Data Structuring
The purpose of this file is to structure the data in a format which makes aggregate and time series analysis and visualization easier. We start with the Cleaned Data `EPL.csv` for this. 

```{r}
EPL <- read.csv("Clean Data/EPL.csv")
colnames(EPL) <- c("Date",colnames(EPL)[-1])
head(EPL)
```
 The structure of the reformatted table will be as follows:
 - Season
 - Year
 - Week
 - Team
 - Opponent
 - Home_Away
 - Goals_Scored
 - Goals_Conceded
 - Points_Earned_Week
 - Points_Total
 - Shots_Taken
 - Shots_Conceded
 - Shots_Taken_Total
 - Shots_Conceded_Total
 - Shots_on_Target
 - Shots_on_Target_Conceded
 - Shots_on_Target_Total
 - Shots_on_Target_Conceded_Total
 - Fouls_Committed
 - Fouls_Against
 - Fouls_Committed_Total
 - Fouls_Against_Total
 - Corners
 - Corners_Conceded
 - Corners_Total
 - Corners_Conceded_Total
 - Yellow_Cards
 - Yellow_Cards_Total
 - Red_Cards
 - Red_Cards_Total
 
 This table will be sorted by Season, Week and Team in that order. 
 
```{r}
EPL.aggregate <- data.frame(matrix(NA,nrow=1,ncol=30))
unique_seasons <- sort(unique(EPL$season))

## Even though, it's a triple loop, we only look at subset of data at each stage hence still O(2*N)
## 2*N because we look at each match twice. 

for(i in seq_along(unique_seasons)){ ## looping through each season
  tmp <- EPL[EPL$season == unique_seasons[i],]
  tmp <- tmp[order(tmp$Date),] ## ensuring sort by date
  unique_teams <- sort(unique(tmp$HomeTeam)) ## looping through teams for the season
  
  season.df <- data.frame(matrix(NA,nrow=1,ncol=30))
  
  for(j in seq_along(unique_teams)){  
    games <- grepl(unique_teams[j],as.character(tmp$HomeTeam)) |
      grepl(unique_teams[j],as.character(tmp$AwayTeam))
    games <- tmp[games,]
    
    team.df <- data.frame(matrix(0,nrow=(dim(games)[1]+1),ncol=30))
    team.df[,1] <- rep(unique_seasons[i],(dim(games)[1]+1))
    team.df[,3] <- c(0:dim(games)[1])
    team.df[,4] <- rep(unique_teams[j],(dim(games)[1]+1))
    
    for(k in c(1:dim(games)[1])){ ## looping through games in season
      team.df[(k+1),2] <- games$year[k]
      h_a <- grep(unique_teams[j],as.matrix(games[k,2:3]))
      team.df[(k+1),5] <- ifelse(h_a == 1,as.character(games[k,3]),as.character(games[k,2]))
      team.df[(k+1),6] <- ifelse(h_a == 1,"H","A")
      team.df[(k+1),7] <- ifelse(h_a == 1,games[k,4],games[k,5])
      team.df[(k+1),8] <- ifelse(h_a == 1,games[k,5],games[k,4])
      team.df[(k+1),9] <- (team.df[(k+1),7] > team.df[(k+1),8])*2 + 1 - (team.df[(k+1),7] < team.df[(k+1),8])*1
      team.df[(k+1),10] <- team.df[(k),10] + team.df[(k+1),9]
      team.df[(k+1),11] <- ifelse(h_a == 1,games[k,11],games[k,12])
      team.df[(k+1),12] <- ifelse(h_a == 1,games[k,12],games[k,11])
      team.df[(k+1),13] <- team.df[(k),13] + team.df[(k+1),11]
      team.df[(k+1),14] <- team.df[(k),14] + team.df[(k+1),12]
      team.df[(k+1),15] <- ifelse(h_a == 1,games[k,13],games[k,14])
      team.df[(k+1),16] <- ifelse(h_a == 1,games[k,14],games[k,13])
      team.df[(k+1),17] <- team.df[(k),17] + team.df[(k+1),15]
      team.df[(k+1),18] <- team.df[(k),18] + team.df[(k+1),16]
      team.df[(k+1),19] <- ifelse(h_a == 1,games[k,15],games[k,16])
      team.df[(k+1),20] <- ifelse(h_a == 1,games[k,16],games[k,15])
      team.df[(k+1),21] <- team.df[(k),21] + team.df[(k+1),19]
      team.df[(k+1),22] <- team.df[(k),22] + team.df[(k+1),20]
      team.df[(k+1),23] <- ifelse(h_a == 1,games[k,17],games[k,18])
      team.df[(k+1),24] <- ifelse(h_a == 1,games[k,18],games[k,17])
      team.df[(k+1),25] <- team.df[(k),25] + team.df[(k+1),23]
      team.df[(k+1),26] <- team.df[(k),26] + team.df[(k+1),24]
      team.df[(k+1),27] <- ifelse(h_a == 1,games[k,19],games[k,20])
      team.df[(k+1),28] <- team.df[(k),28] + team.df[(k+1),27]
      team.df[(k+1),29] <- ifelse(h_a == 1,games[k,21],games[k,22])
      team.df[(k+1),30] <- team.df[(k),30] + team.df[(k+1),29]
    }
    team.df <- team.df[-1,]
    season.df <- rbind(season.df,team.df)
  }
  season.df <- season.df[-1,]
  season.df <- season.df[order(season.df[,3]),]
  EPL.aggregate <- rbind(EPL.aggregate,season.df)
}

EPL.aggregate <- EPL.aggregate[-1,]
colnames(EPL.aggregate) <- c("Season","Year","Week","Team","Opponent","Home_Away","Goals_Scored",
                             "Goals_Conceded","Points_Earned_Week","Points_Total","Shots_Taken",
                             "Shots_Conceded","Shots_Taken_Total","Shots_Conceded_Total","Shots_on_Target",
                             "Shots_on_Target_Conceded","Shots_on_Target_Total",
                             "Shots_on_Target_Conceded_Total",
                             "Fouls_Committed","Fouls_Against","Fouls_Committed_Total","Fouls_Against_Total",
                             "Corners","Corners_Conceded","Corners_Total","Corners_Conceded_Total",
                             "Yellow_Cards","Yellow_Cards_Total","Red_Cards","Red_Cards_Total")


write.csv(EPL.aggregate,file="Clean Data/EPL_aggregate.csv",row.names = FALSE)

head(EPL.aggregate)


```
 
 